The CrypTool Script: \\ Cryptography, Mathematics and more \\ \vspace{30pt} {\large Background reading\\%*[-8pt] for the free eLearning program CrypTool\\ (8th edition -- distributed with CrypTool version 1.4.10)} \vspace{60pt}

### 2007-07-02 script/de/chapters/primes.tex
Update scripts: chap 8, appendix a.4, quotes, minor corrections, minor layout improvements

% --------------------------------------------------------------------------
\subsection{Primzahltests}
\label{primality_tests}   % chap. 3.5
\index{Primzahltest}

@@

% --------------------------------------------------------------------------
% \pagebreak %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%   be_2005
\vskip +10 pt
\subsubsection{$f(n) = n^2 - 79 \cdot n + 1.601$}
\label{L-Polynomfunktion02-1601}
   Diese Funktion liefert für die Werte $n=0$ bis $n=79$ stets Primzahlwerte.
   Leider ergibt $f(80) = 1.681 = 11 \cdot 151$ keine Primzahl. Bis heute
   kennt man keine Funktion, die mehr aufeinanderfolgende Primzahlen annimmt.
   Andererseits kommt jede Primzahl doppelt vor (erst in der absteigenden,
   dann in der aufsteigenden Folge), so dass sie insgesamt genau 40 
   verschiedene Primzahlwerte liefert (dieselben wie die, die die Funktion
   aus Kapitel~\ref{L-Polynomfunktion01-41} liefert).
$$
\begin{array}{|ll||ll|}
\hline
f(0) = 1.601    & \mapsto {\rm ~prim} &  f(28) = 173    & \mapsto {\rm ~prim} \\
f(1) = 1.523    & \mapsto {\rm ~prim} &  f(29) = 151    & \mapsto {\rm ~prim} \\
f(2) = 1.447    & \mapsto {\rm ~prim} &  f(30) = 131 & \mapsto {\rm ~prim} \\
f(3) = 1.373    & \mapsto {\rm ~prim} &  f(31) = 113 & \mapsto {\rm ~prim} \\
f(4) = 1.301    & \mapsto {\rm ~prim} &  f(32) = 97 & \mapsto {\rm ~prim} \\
f(5) = 1.231    & \mapsto {\rm ~prim} &  f(33) = 83 & \mapsto {\rm ~prim} \\
f(6) = 1.163    & \mapsto {\rm ~prim} &  f(34) = 71 & \mapsto {\rm ~prim} \\
f(7) = 1.097    & \mapsto {\rm ~prim} &  f(35) = 61 & \mapsto {\rm ~prim} \\
f(8) = 1.033    & \mapsto {\rm ~prim} &  f(36) = 53 & \mapsto {\rm ~prim} \\
f(9) = 971  & \mapsto {\rm ~prim} &  f(37) = 47 & \mapsto {\rm ~prim} \\
f(10) = 911 & \mapsto {\rm ~prim} &  f(38) = 43 & \mapsto {\rm ~prim} \\
f(11) = 853 & \mapsto {\rm ~prim} &  f(39) = 41 & \mapsto {\rm ~prim} \\
f(12) = 797 & \mapsto {\rm ~prim} &  f(40) = 41 & \mapsto {\rm ~prim} \\                 
f(13) = 743 & \mapsto {\rm ~prim} &  f(41) = 43 & \mapsto {\rm ~prim} \\                 
f(14) = 691 & \mapsto {\rm ~prim} &  f(42) = 47 & \mapsto {\rm ~prim} \\                 
f(15) = 641 & \mapsto {\rm ~prim} &  f(43) = 53 & \mapsto {\rm ~prim} \\                 
f(16) = 593 & \mapsto {\rm ~prim} &  \cdots  &  \\                                       
f(17) = 547 & \mapsto {\rm ~prim} &  f(77) = 1.447  & \mapsto {\rm ~prim} \\             
f(18) = 503 & \mapsto {\rm ~prim} &  f(78) = 1.523  & \mapsto {\rm ~prim} \\             
f(19) = 461 & \mapsto {\rm ~prim} &  f(79) = 1.601  & \mapsto {\rm ~prim} \\             
f(20) = 421 & \mapsto {\rm ~prim} &  f(80) = 41 \cdot 41 & \mapsto {\rm ~NICHT~prim!} \\
f(21) = 383 & \mapsto {\rm ~prim} &  f(81) = 41 \cdot 43 & \mapsto {\rm ~NICHT~prim!} \\ 
f(22) = 347 & \mapsto {\rm ~prim} &  f(82) = 1.847  & \mapsto {\rm ~prim} \\             
f(21) = 383 & \mapsto {\rm ~prim} &  f(83) = 1.933  & \mapsto {\rm ~prim} \\             
f(22) = 347 & \mapsto {\rm ~prim} &  f(84) = 43 \cdot 47 &  \mapsto {\rm ~NICHT~prim!} \\
f(23) = 313 & \mapsto {\rm ~prim} & & \\ 
f(24) = 281 & \mapsto {\rm ~prim} & & \\
f(25) = 251 & \mapsto {\rm ~prim} & & \\
f(26) = 223 & \mapsto {\rm ~prim} & & \\
f(27) = 197 & \mapsto {\rm ~prim} & & \\
\hline
\end{array}
$$

@@

Ein einfacher Weg, alle $PI(x)$ Primzahlen kleiner oder gleich $x$
zu berechnen, ist das Sieb des Erathostenes. Er fand schon im 3.
Jahrhundert vor Christus einen sehr einfach automatisierbaren Weg,
das herauszufinden. Zuerst werden ab 2 alle Zahlen bis $x$
aufgeschrieben, die 2 umkreist und dann streicht man alle
Vielfachen von 2. Anschließend umkreist man die kleinste noch
nicht umkreiste oder gestrichene Zahl (3), streicht wieder alle
ihre Vielfachen, usw. Durchzuführen braucht man das nur bis zu
der größten Zahl, deren Quadrat kleiner oder gleich $x$ 
ist.\index{ZT, Lernprogramm Zahlentheorie}%
\footnote{%
Mit dem Lernprogramm {\bf ZT} können Sie für beliebige eigene Wertemengen
das Sieb des Eratosthenes rechnergestützt und geführt Schritt für Schritt
anwenden: Siehe Lern-Kapitel 1.2, Seiten 6/21 und 7/21.\\
ZT können Sie in CrypTool\index{CrypTool} über das Menü
{\bf Einzelverfahren \textbackslash{} Zahlentheorie interaktiv \textbackslash{}
Lernprogramm für Zahlentheorie} aufrufen.
}

@@

\bibitem[Aaronson2003]{Aaronson2003} \index{Aaronson 2003}
    Scott Aaronson, \\
    {\em The Prime Facts: From Euclid to AKS}, \\
    \href{http://www.cs.berkeley.edu/~aaronson/prime.ps}
         {\texttt{http://www.cs.berkeley.edu/\~{}aaronson/prime.ps}}. \\
    Dieses schöne Paper wurde mir erst nach Vollendung dieses Artikels
    bekannt: es bietet ebenfalls einen einfachen, didaktisch 
    gut aufbereiteten und trotzdem nicht oberflächlichen Einstieg 
    in dieses Thema (ist leider nur in Englisch verfügbar).

@@

\vskip +10 pt
% --------------------------------------------------------------------------(5\cdot2^7 - 1)(5^2\cdot2^14 + 1)
\subsection*{Dank} \addcontentsline{toc}{subsection}{Dank}
Für das sehr konstruktive Korrekturlesen dieses Artikels: Hr.
Henrik Koy und Hr. Roger Oyono.

>>>>

% --------------------------------------------------------------------------
\subsection[Primzahltests]{Primzahltests\footnotemark}
\footnotetext{%
    \index{ZT, Lernprogramm Zahlentheorie}%
    \index{Lernprogramm ZT}%
    In dem Lernprogramm {\bf ZT} können Sie den Fermat-Test und den Miller-Rabin-Test
    geführt Schritt für Schritt anwenden: Siehe Lern-Kapitel 3.2 und 3.3, 
    Seiten 3-11/11.\\
    ZT können Sie in CrypTool\index{CrypTool} über das Menü
    {\bf Einzelverfahren \textbackslash{} Zahlentheorie interaktiv
    \textbackslash{} Lernprogramm für Zahlentheorie} aufrufen.
    Siehe Anhang \ref{s:appendix-Learn-NT}.
}
\label{primality_tests}   % chap. 3.5
\index{Primzahltest}

@@

% --------------------------------------------------------------------------
% \pagebreak %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%   be_2005
\vskip +10 pt
\subsubsection{$f(n) = n^2 - 79 \cdot n + 1.601$}
\label{L-Polynomfunktion02-1601}   \label{SrcPrimArith1}
   Diese Funktion liefert für die Werte $n=0$ bis $n=79$ stets
   Primzahlwerte\footnote{%
   In \hyperlink{PrimArith1}{Anhang A zu diesem Kapitel} finden Sie den
   Quellcode zur Berechnung der Tabelle mit Pari-GP.\index{Pari-GP}}.
   Leider ergibt $f(80) = 1.681 = 11 \cdot 151$ keine Primzahl. Bis heute
   kennt man keine Funktion, die mehr aufeinanderfolgende Primzahlen annimmt.
   Andererseits kommt jede Primzahl doppelt vor (erst in der absteigenden,
   dann in der aufsteigenden Folge), so dass sie insgesamt genau 40 
   verschiedene Primzahlwerte in Folge liefert (Es sind dieselben wie die, die
   die Funktion aus Kapitel~\ref{L-Polynomfunktion01-41} liefert)\footnote{%
   Ein weiteres quadratisches Polynom, das dieselben Primzahlen liefert, ist: 
   $f(n) = n^2 - 9 \cdot n + 61$.\\ Unter den ersten 1000 Folgengliedern dieser
   Funktion sind über 50\% prim (vgl. \hyperlink{PrimArith2}{Anhang A zu diesem
   Kapitel}).}.
$$
\begin{array}{| ll       || ll |}
\hline
f(0)  = 1.601 & \mapsto {\rm ~prim}   &   f(26) =   223 & \mapsto {\rm ~prim} \\
f(1)  = 1.523 & \mapsto {\rm ~prim}   &   f(27) =   197 & \mapsto {\rm ~prim} \\
f(2)  = 1.447 & \mapsto {\rm ~prim}   &   f(28) =   173 & \mapsto {\rm ~prim} \\
f(3)  = 1.373 & \mapsto {\rm ~prim}   &   f(29) =   151 & \mapsto {\rm ~prim} \\
f(4)  = 1.301 & \mapsto {\rm ~prim}   &   f(30) =   131 & \mapsto {\rm ~prim} \\
f(5)  = 1.231 & \mapsto {\rm ~prim}   &   f(31) =   113 & \mapsto {\rm ~prim} \\
f(6)  = 1.163 & \mapsto {\rm ~prim}   &   f(32) =    97 & \mapsto {\rm ~prim} \\
f(7)  = 1.097 & \mapsto {\rm ~prim}   &   f(33) =    83 & \mapsto {\rm ~prim} \\
f(8)  = 1.033 & \mapsto {\rm ~prim}   &   f(34) =    71 & \mapsto {\rm ~prim} \\
f(9)  =   971 & \mapsto {\rm ~prim}   &   f(35) =    61 & \mapsto {\rm ~prim} \\
f(10) =   911 & \mapsto {\rm ~prim}   &   f(36) =    53 & \mapsto {\rm ~prim} \\
f(11) =   853 & \mapsto {\rm ~prim}   &   f(37) =    47 & \mapsto {\rm ~prim} \\
f(12) =   797 & \mapsto {\rm ~prim}   &   f(38) =    43 & \mapsto {\rm ~prim} \\
f(13) =   743 & \mapsto {\rm ~prim}   &   f(39) =    41 & \mapsto {\rm ~prim} \\
f(14) =   691 & \mapsto {\rm ~prim}   &   f(40) =    41 & \mapsto {\rm ~prim} \\              
f(15) =   641 & \mapsto {\rm ~prim}   &   f(41) =    43 & \mapsto {\rm ~prim} \\              
f(16) =   593 & \mapsto {\rm ~prim}   &   f(42) =    47 & \mapsto {\rm ~prim} \\              
f(17) =   547 & \mapsto {\rm ~prim}   &   f(43) =    53 & \mapsto {\rm ~prim} \\              
f(18) =   503 & \mapsto {\rm ~prim}   &   \cdots  &  \\                                       
f(19) =   461 & \mapsto {\rm ~prim}   &   f(77) = 1.447 & \mapsto {\rm ~prim} \\             
f(20) =   421 & \mapsto {\rm ~prim}   &   f(78) = 1.523 & \mapsto {\rm ~prim} \\             
f(21) =   383 & \mapsto {\rm ~prim}   &   f(79) = 1.601 & \mapsto {\rm ~prim} \\             
f(22) =   347 & \mapsto {\rm ~prim}   &   f(80) = 41 \cdot 41 & \mapsto {\rm ~NICHT~prim!} \\
f(21) =   383 & \mapsto {\rm ~prim}   &   f(81) = 41 \cdot 43 & \mapsto {\rm ~NICHT~prim!} \\ 
f(22) =   347 & \mapsto {\rm ~prim}   &   f(82) = 1.847 & \mapsto {\rm ~prim} \\             
f(23) =   313 & \mapsto {\rm ~prim}   &   f(83) = 1.933 & \mapsto {\rm ~prim} \\              
f(24) =   281 & \mapsto {\rm ~prim}   &   f(84) = 43 \cdot 47 &  \mapsto {\rm ~NICHT~prim!} \\
f(25) =   251 & \mapsto {\rm ~prim}   &   & \\
\hline
\end{array}
$$

@@

Ein einfacher Weg, alle $PI(x)$ Primzahlen kleiner oder gleich $x$
zu berechnen, ist das Sieb des Erathostenes. Er fand schon im 3.
Jahrhundert vor Christus einen sehr einfach automatisierbaren Weg,
das herauszufinden. Zuerst werden ab 2 alle Zahlen bis $x$
aufgeschrieben, die 2 umkreist und dann streicht man alle
Vielfachen von 2. Anschließend umkreist man die kleinste noch
nicht umkreiste oder gestrichene Zahl (3), streicht wieder alle
ihre Vielfachen, usw. Durchzuführen braucht man das nur bis zu
der größten Zahl, deren Quadrat kleiner oder gleich $x$ ist.%
\footnote{%
\index{ZT, Lernprogramm Zahlentheorie}%
\index{Lernprogramm ZT}%
Mit dem Lernprogramm {\bf ZT} können Sie für beliebige eigene Wertemengen
das Sieb des Eratosthenes rechnergestützt und geführt Schritt für
Schritt anwenden: Siehe Lern-Kapitel 1.2, Seite 6/21 und 7/21.\\
ZT können Sie in CrypTool\index{CrypTool} über das Menü
{\bf Einzelverfahren \textbackslash{} Zahlentheorie interaktiv \textbackslash{}
Lernprogramm für Zahlentheorie} aufrufen.
    Siehe Anhang \ref{s:appendix-Learn-NT}.}

@@

\bibitem[Aaronson2003]{Aaronson2003} \index{Aaronson 2003}
    Scott Aaronson, \\
    {\em The Prime Facts: From Euclid to AKS}, \\
    % veraltete Seite: http://www.cs.berkeley.edu/~aaronson/prime.ps
    \href{http://www.scottaaronson.com/writings/prime.pdf}
 {\texttt{http://www.scottaaronson.com/writings/prime.pdf}}. \\
    Dieses schöne Paper wurde mir erst nach Vollendung dieses Artikels
    bekannt: Es bietet ebenfalls einen einfachen, didaktisch 
    gut aufbereiteten und trotzdem nicht oberflächlichen Einstieg 
    in dieses Thema (ist nur in Englisch verfügbar).

@@

\vskip +10 pt
% --------------------------------------------------------------------------(5\cdot2^7 - 1)(5^2\cdot2^14 + 1)
\subsection*{Dank} \addcontentsline{toc}{subsection}{Dank}
Für das sehr konstruktive Korrekturlesen der ersten Versionen
dieses Artikels: Hr. Henrik Koy und Hr. Roger Oyono.

% ++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\newpage
\hypertarget{Appendix_A}{}
\subsection*{Anhang A: Beispiele mit Pari-GP }
\addcontentsline{toc}{subsection}{Anhang A: Beispiele mit Pari-GP}
\index{Pari-GP}

In diesem Anhang finden Sie den Quellcode für Pari-GP, mit dem man
z.B. die quadratischen Beispielfolgen für Primzahlen berechnen kann.

\hypertarget{PrimArith1}{}
\subsubsection*{Bewertung des Ergebnisses einer quadratischen \glqq Primzahlformel''}
Folgen wie die in der auf Seite \pageref{SrcPrimArith1} aufgeführten
Funktionswertetabelle können analog der folgenden Pari-GP-Anweisungen
erstellt werden:
\vspace{-2ex}
\begin{verbatim}
print("N -- N^2 - 9*N + 61");
for(N=0,50, X=N^2 - 9*N + 61;
  if (isprime(X)==1,
    print(N, " -- ", X, "  is prime"),
    print(N, " -- ", X, "  is NOT prime")
     );
   );
\end{verbatim}
Der Output sieht folgendermaßen aus:
\vspace{-2ex}
\begin{verbatim}
N -- N^2 - 9*N + 61
0 -- 61  is prime
1 -- 53  is prime
2 -- 47  is prime
3 -- 43  is prime
4 -- 41  is prime
5 -- 41  is prime
6 -- 43  is prime
7 -- 47  is prime
8 -- 53  is prime
9 -- 61  is prime
10 -- 71  is prime
11 -- 83  is prime
12 -- 97  is prime
13 -- 113  is prime
14 -- 131  is prime
15 -- 151  is prime
16 -- 173  is prime
17 -- 197  is prime
18 -- 223  is prime
19 -- 251  is prime
20 -- 281  is prime
21 -- 313  is prime
22 -- 347  is prime
23 -- 383  is prime
24 -- 421  is prime
25 -- 461  is prime
26 -- 503  is prime
27 -- 547  is prime
28 -- 593  is prime
29 -- 641  is prime
30 -- 691  is prime
31 -- 743  is prime
32 -- 797  is prime
33 -- 853  is prime
34 -- 911  is prime
35 -- 971  is prime
36 -- 1033  is prime
37 -- 1097  is prime
38 -- 1163  is prime
39 -- 1231  is prime
40 -- 1301  is prime
41 -- 1373  is prime
42 -- 1447  is prime
43 -- 1523  is prime
44 -- 1601  is prime
45 -- 1681  is NOT prime
46 -- 1763  is NOT prime
47 -- 1847  is prime
48 -- 1933  is prime
49 -- 2021  is NOT prime
50 -- 2111  is prime
\end{verbatim}

\hypertarget{PrimArith2}{}
\subsubsection*{Prozentsatz an Primzahlen, die die quadratische \glqq Primzahlformel'' liefert}

Pari-GP-Anweisung:
\vspace{-2ex}
\begin{verbatim}
A=0;Beg=0;End=1000;
for(N=Beg,End, X=N^2-9*N+61;
  if(isprime(X)==1, 
  print(N, " -- ", X, "  is prime");A++,
  print(N, " -- ", X, "  is NOT prime")
    ) );
print("Number of primes: ", A, ";  Percentage of primes: " (A*100)/((End-Beg)+1)*1.);
\end{verbatim}
Der Output sieht folgendermaßen aus:
\vspace{-2ex}
\begin{verbatim}
...
1000 -- 991061  is NOT prime
Number of primes: 584;  Percentage of primes: 58.34165834165834165834165834
\end{verbatim}

### 2007-07-02 script/en/chapters/authors.tex
Update scripts: chap 8, appendix a.4, quotes, minor corrections, minor layout improvements

@@

Cryptographic applications -- classical as well as modern ones -- have been
used in literature and movies. In some media they are only mentioned and
are a pure addmixture; in others they play a primary role and are explained
in detail; and sometimes the purpose of the story, which forms the framework,
is primarily to transport this knowledge and achieve better motivation.

Here is the beginning of an overview:

@@

\vskip +60 pt
Kid books with collections of simpler cryptographic encryption methods, 
prepared in a didactical and exciting manner are in the following list
(please send us similar English kid books, because at the moment
our list contains only German kid books):

@@

\clearpage
\listoffigures
\addcontentsline{toc}{section}{\listfigurename}

\clearpage

\listoftables
\addcontentsline{toc}{section}{\listtablename}

>>>>

---------

\item[Johannes Buchmann,] \mbox{}\\
Co-author of the chapter ``Crypto2020 --- Perspectives for long-term cryptographic security''.
%Vice-president of the Technische Universit\"at Darmstadt, Germany and
%Professor at the Departments of Computer Science and of Mathematics.
Johannes Buchmann holds the Chair for Theoretical Computer Science (Cryptography
and Computer Algebra) at the department of Computer Science of the Technische
Universit\"at Darmstadt TUD).  He is also a Professor at the department of Mathematics, and vice-president of the university.

\item[Alexander May,] \mbox{}\\
Co-author of the chapter ``Crypto2020 --- Perspectives for long-term cryptographic security'', 
Assistant Professor at the department of Computer Science of
the Technische Universit\"at Darmstadt, Germany.

\item[Erik Dahmen,] \mbox{}\\
Co-author of the chapter ``Crypto2020 --- Perspectives for long-term cryptographic security'', 
%Researcher at the Department of Computer Science of the TUD.
Researcher at the Chair for Theoretical Computer Science (Cryptography and
Computer Algebra), department of Computer Science, Technische Universit\"at
Darmstadt, Germany.

\item[Ulrich Vollmer,] \mbox{}\\
Co-author of the chapter ``Crypto2020 --- Perspectives for long-term cryptographic security'', 
Researcher at the Chair for Theoretical Computer Science (Cryptography and
Computer Algebra), department of Computer Science, Technische Universit\"at
Darmstadt, Germany.

@@

Cryptographic applications -- classical as well as modern ones -- have been
used in literature and movies. In some media they are only mentioned and
are a pure addmixture; in others they play a primary role and are explained
in detail; and sometimes the purpose of the story, which forms the framework,
is primarily to transport this knowledge and achieve better motivation.
Here is the beginning of an overview.

\paragraph{For grownups:}

@@

\vskip +60 pt
\paragraph{For kids:}

Kid books with collections of simpler cryptographic encryption methods, 
prepared in a didactical and exciting manner are in the following list
(please send us similar English kid books, because at the moment
our list contains only German kid books):

@@

% ++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\newpage
\hypertarget{appendix-Learn-NT}{}
\subsection{Learning Tool for Elementary Number Theory}
    \label{s:appendix-Learn-NT}  
    \index{NT, Learning Tool for Number Theory}%
    \index{Educational tool NT}%

CrypTool contains an interactive educational tool for elementary number theory,
called ``NT''.\footnote{%
    NT can be called in CrypTool\index{CrypTool} via the menu path
    {\bf Indiv. Procedures \textbackslash{} Number Theory Interactive
    \textbackslash{} Learning Tool for Number Theory}.
}

The educational tool ``NT'' (nunber theory) by Martin Ramberger introduces
number theory and visualizes many of the methods and concepts.
Where necessary it show the according mathematical formulas.
Often you can apply the mathematical methods dynamically with your own small
numerical examples.

The content of this educational tool is mainly based on the books 
\cite{Buchmann2004} and \cite{Scheid2003}.

This visualized educational tool was build with Authorware 4.

\paragraph{Request for znhancement/zpgrade:}
It would be desirable to update it to a new version Authorware or to use another
development platform. If there are developers interested to do this, I'd be
more than happy (please send an email at the author of this CrypTool script).

\paragraph{Figures:}
The figures ~\ref{NT_Fig_C1.3_EuclidsAlg-LinearCombinations}
till~\ref{NT_Fig_C5.3_PollardRho} give you an impression of the
educational tool ``NT'':

% -> Figur 1  WIE ERREICHT MAN, dass er von vorne zählt ? (nicht wichtig)
\begin{figure}[h]
\begin{center}
\includegraphics[scale=0.4]{figures/NT_Fig_C1-3_EuclidsAlg-LinearCombinations.png}
\caption{Each common divisor of two integers also divides all its linear
combinations\vspace{1ex}} 
\label{NT_Fig_C1.3_EuclidsAlg-LinearCombinations}
\end{center}
\end{figure}

\begin{figure}[h]
\begin{center}
\includegraphics[scale=0.4]{figures/NT_Fig_C1-3_EuclidsAlg-Procedure}
\caption{Euklid's algorithm to determine gcd\vspace{1ex}} 
\label{NT_Fig_C1.3_EuclidsAlg-Procedure}
\end{center}
\end{figure}

\begin{figure}[h]
\begin{center}
\includegraphics[scale=0.4]{figures/NT_Fig_C3-1_PrimesDistribution}
\caption{Distribution of primes and its differences\vspace{1ex}} 
\label{NT_Fig_C3.1_PrimesDistribution}
\end{center}
\end{figure}

\begin{figure}[h]
\begin{center}
\includegraphics[scale=0.4]{figures/NT_Fig_C3-2_Fermat-Test}
\caption{Finding primes with the prime number test of Fermat\vspace{1ex}} 
\label{NT_Fig_C3.2_Fermat-Test}
\end{center}
\end{figure}

\begin{figure}[h]
\begin{center}
\includegraphics[scale=0.4]{figures/NT_Fig_C4-1_ReversibilityAdditiveCipher}
\caption{Reversibility of encryption mechanisms exemplified with additive
ciphers\vspace{1ex}} 
\label{NT_Fig_C4.1_ReversibilityAdditiveCipher}
\end{center}
\end{figure}

\begin{figure}[h]
\begin{center}
\includegraphics[scale=0.4]{figures/NT_Fig_C5-2_Fermat-factorization-How-far-1}
\caption{Fermat factorisation using the third binomial formula\vspace{1ex}} 
\label{NT_Fig_C5.2_Fermat-factorization-How-far-1}
\end{center}
\end{figure}

\begin{figure}[h]
\begin{center}
\includegraphics[scale=0.4]{figures/NT_Fig_C5-2_Fermat-factorization-How-far-2}
\caption{Fermat factorisation: Detecting squares\vspace{1ex}} 
\label{NT_Fig_C5.2_Fermat-factorization-How-far-2}
\end{center}
\end{figure}

\begin{figure}[h]
\begin{center}
\includegraphics[scale=0.4]{figures/NT_Fig_C5-3_PollardRho}
\caption{Pollard's rho factorisation: Floyd's cycle finding algorithm\vspace{1ex}} 
\label{NT_Fig_C5.3_PollardRho}
\end{center}
\end{figure}

% ++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\clearpage
%\newpage
\begin{thebibliography}{99999}
%\addcontentsline{toc}{subsection}{Literaturverzeichnis}

\bibitem[Buchmann2004]{Buchmann2004} \index{Buchmann 2004}
    Johannes Buchmann, \\
    {\em Introduction to Cryptography}, Springer, 2nd edition, 2004.

\bibitem[Scheid2003]{Scheid2003} \index{Scheid 2003}
    Harald Scheid, \\
    {\em Zahlentheorie}, Spektrum Akademischer Verlag, 3rd edition, 2003.

\end{thebibliography}

% ++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\clearpage
\listoffigures
\addcontentsline{toc}{section}{\listfigurename}

\clearpage
\listoftables
\addcontentsline{toc}{section}{\listtablename}

### 2007-07-02 script/en/chapters/crypto2020.tex
Update scripts: chap 8, appendix a.4, quotes, minor corrections, minor layout improvements

In 2005, that is 27 years after the invention of
RSA, Bahr, Boehm, Franke, and Kleinjung from
Bochum University managed to factor a 200 digit
RSA challenge number
(\url{www.mat.uniroma2.it/~eal/rsa640.txt}).  A
key with size originally thought to be secure for
a very long time was broken with a computation
that took them just five months.  This illustrates
the tremendous progress factoring technology has
made within the last 30 years.  This progress is
based on break-through mathematical ideas --- e.g.\
the number field sieve proposed by John
Pollard --- as well as significant developments in
computer hardware and software implementation
technology.\footnote{%
Please compare chapter \ref{SecurityRSA}
\hyperlink{SecurityRSA}{Considerations regarding the
security of the RSA algorithm}, and especailly chapters
\ref{NoteFactorisation} and \ref{FactorisationResearch}.
}

@@

Yet, even a well thought out extrapolation formula
is no security guarantee!  At any time, a
brilliant mathematical idea can allow us to factor
large numbers easily, and destroy the security of
RSA.  In 1996, Peter Shor showed that a quantum
computer --- a new type of computer that leverages
the laws of quantum mechanics to speed up certain
types of computations --- can in principle be used
for the fast factorization of large numbers \cite{shor:1997}.
Despite intensive research in the area, it is
still too early to judge whether we are ever going
to be able to build quantum computers\index{Quantum computer} of
sufficient capacity to apply Shor's algorithm to
numbers of relevant size.\footnote{%
Required qbits for attacks on RSA, DSA and ECDSA using key with a bit length n: \\
\vskip +1 pt
\begin{tabular}{|c|l|}
\hline
   RSA		&  2n + 3 \\
   DSA		&  2n + 3 \\
   ECDSA $2^n$	&  \~{}2n + 8 log n \\
   ECDSA p	&  \~{}6n \\
\hline
\end{tabular}
\vskip +6 pt
Please compare chapter 5.3 in
``SicAri -- Eine Sicherheitsplattform und deren Werkzeuge
für die ubiquitäre Internetnutzung, KB2.1 -- Abschlussbericht,
Übersicht über Angriffe auf relevante kryptographische Verfahren'',
version 1.0, Mai 17, 2005,
Prof. Dr. Johannes Buchmann et al., TUD-KryptC and cv cryptovision GmbH
(\href{http://www.cdc.informatik.tu-darmstadt.de/~schepers/kb\_21\_angriffe.pdf}
 {\tt http://www.cdc.informatik.tu-darmstadt.de/\~{}schepers/kb\_21\_angriffe.pdf}).
}
Recent announcements of
significant progress in this area made by the
start-up company D-Wave (\url{www.dwavesys.com})
have been greeted with a lot of scepticism, even
ridicule.

@@

\subsection{Quantum cryptography -- a loophole?}
\label{sec:quantum cryptography}\index{Quantum cryptography}

>>>>

In 2005, that is 27 years after the invention of
RSA, Bahr, Boehm, Franke, and Kleinjung from
Bochum University managed to factor a 200 digit
RSA challenge number
(\url{www.mat.uniroma2.it/~eal/rsa640.txt}).  A
key with size originally thought to be secure for
a very long time was broken with a computation
that took them just five months.  This illustrates
the tremendous progress factoring technology has
made within the last 30 years.  This progress is
based on break-through mathematical ideas --- e.g.\
the number field sieve proposed by John
Pollard --- as well as significant developments in
computer hardware and software implementation
technology.\footnote{%
Please compare chapter \ref{SecurityRSA}
\hyperlink{SecurityRSA}{Considerations regarding the
security of the RSA algorithm}, and especially chapters
\ref{NoteFactorisation} and \ref{FactorisationResearch}.
}

@@

Yet, even a well thought out extrapolation formula
is no security guarantee!  At any time, a
brilliant mathematical idea can allow us to factor
large numbers easily, and destroy the security of
RSA.  In 1996, Peter Shor showed that a quantum
computer --- a new type of computer that leverages
the laws of quantum mechanics to speed up certain
types of computations --- can in principle be used
for the fast factorization of large numbers \cite{shor:1997}.
Despite intensive research in the area, it is
still too early to judge whether we are ever going
to be able to build quantum computers\index{Quantum computer} of
sufficient capacity to apply Shor's algorithm to
numbers of relevant size.\footnote{%
Required qbits for attacks on RSA, DSA and ECDSA using key with a bit length n: \\
\vskip +1 pt
\begin{tabular}{|c|l|}
\hline
   RSA		&  2n + 3 \\
   DSA		&  2n + 3 \\
   ECDSA $2^n$	&  \~{}2n + 8 log n \\
   ECDSA p	&  \~{}4n \\
\hline
\end{tabular}
\vskip +6 pt
Please compare chapter 5.3 in
``SicAri -- Eine Sicherheitsplattform und deren Werkzeuge
f\"ur die ubiquit\"are Internetnutzung, KB2.1 -- Abschlussbericht,
\"Ubersicht \"uber Angriffe auf relevante kryptographische Verfahren'',
version 1.0, Mai 17, 2005,
Prof. Dr. Johannes Buchmann et al., TUD-KryptC and cv cryptovision GmbH
(\href{http://www.cdc.informatik.tu-darmstadt.de/~schepers/kb\_21\_angriffe.pdf}
 {\tt http://www.cdc.informatik.tu-darmstadt.de/\~{}schepers/kb\_21\_angriffe.pdf}) and the dissertation of Axel Schmidt at the same faculty.
}
Recent announcements of
significant progress in this area made by the
start-up company D-Wave (\url{www.dwavesys.com})
have been greeted with a lot of scepticism, even
ridicule.

@@

%\subsection{Quantum cryptography -- a loophole?}
\subsection{Quantum cryptography -- a way out of the impasse?}
\label{sec:quantum cryptography}\index{Quantum cryptography}

### 2007-07-02 script/en/chapters/digitalsignatures.tex
Update scripts: chap 8, appendix a.4, quotes, minor corrections, minor layout improvements

An asymmetric technique is used again (see encryption procedures). Participants
who wish to generate a digital signature for a document must possess a pair of
keys. They use their secret key to generate signatures and the recipient uses
the sender's public key to verify whether the signature is correct. As before,
it must be impossible to use the public key to derive the secret key\footnote{%
With CrypTool\index{CrypTool} you can also generate and check 
digital signatures: \\
using the submenus of the main menu {\bf Digital Signatures / PKI} or \\
using menu {\bf Indiv. Procedures \textbackslash{} RSA Cryptosystem 
\textbackslash{} Signature Demonstration (Signature Generation)}.
}.

@@

% --------------------------------------------------------------------------
\vskip + 15pt
\subsubsection{Signing with hash functions}
The hash function procedure is as follows:\\ Rather than signing the actual
document, the sender now first calculates the hash value of the message and
signs this. The recipient also calculates the hash value of the message (the
algorithm used must be known), then verifies whether the signature sent with the
message is a correct signature of the hash value. If this is the case, the
signature is verified to be correct. This means that the message is authentic,
because we have assumed that knowledge of the public key does not enable you to
derive the secret key. However, you would need this secret key to sign messages
in another name.

>>>>

An asymmetric technique is used again (see encryption procedures). Participants
who wish to generate a digital signature for a document must possess a pair of
keys. They use their secret key to generate signatures and the recipient uses
the sender's public key to verify whether the signature is correct. As before,
it must be impossible to use the public key to derive the secret key\footnote{%
With CrypTool\index{CrypTool} you can also generate and check 
digital signatures: Using\\
the submenus of the main menu {\bf Digital Signatures / PKI}  or using\\
menu {\bf Indiv. Procedures \textbackslash{} RSA Cryptosystem 
\textbackslash{} Signature Demonstration (Signature Generation)}.
}.

@@

% --------------------------------------------------------------------------
\vskip + 15pt
\subsubsection{Signing with hash functions}
The signature procedure with hash functions is as follows:\footnote{% 
Compare: \href{http://en.wikipedia.org/wiki/Digital\_signature}
      {\texttt{http://en.wikipedia.org/wiki/Digital\_signature}}.
} \\ Rather than signing the actual
document, the sender now first calculates the hash value of the message and
signs this. The recipient also calculates the hash value of the message (the
algorithm used must be known), then verifies whether the signature sent with the
message is a correct signature of the hash value. If this is the case, the
signature is verified to be correct. This means that the message is authentic,
because we have assumed that knowledge of the public key does not enable you to
derive the secret key. However, you would need this secret key to sign messages
in another name.
